---
title: "Visium Spatial Array and Image Geometry"
subtitle: "Manual Alignment-derived spatial and image information"
author: "Eli Spies"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    anchor_sections: false
    fig_caption: true
    code_folding: hide
    theme: cosmo
    fontsize: 1.1em
    css: css_spatial_resolution.css
    includes:
      after_body: darkmode.html
---


```{r setup, include=FALSE}
# to be added to yaml header
# self_contained: true

# code folding
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center")
```


---

# Motivation


The coordinate transform used for fiducial registration is conceptually part of Space Ranger's alignment procedure, but it is not exposed as a documented field in the standard spatial output files. The Loupe manual-alignment JSON, available after [Manual Alignment](https://www.10xgenomics.com/support/software/space-ranger/latest/analysis/inputs/image-fiducial-alignment) using the 10X Loupe Browser, contains sufficient information to reconstruct a spatial transformation grid. However, 10x Genomics documents that JSON format as internal and not fixed.

Distances in json file and information by Visium matches case 3 geometry from the Reconstruction models as well as the provided spatial information of 100µm distances to nearest neighbored spots remarkably well. 
But, though strong evidence for selecting case 3 as underlying spatial grid, the geometry applied in case 3 fails tests for physical containment and non-overlap for the 6.5 mm and 11 mm Visium arrays, implying that the exact capture area boundary won't enable capturing all spots with distances as given. 

Without further information, the question what geometrical boundary the manufacturer's nominal 6.5-mm and 11-mm capture-area dimensions actually refer to, remains unresolved. Given the strong evidence, we chose case 3 as default spatial grid for further spatially resolved analyses, though.

---
  
# Data availability

The spatial data for three 11mm and three 6.5mm arrays was downloaded from the 
[10X Genomics Datasets download site](https://www.10xgenomics.com/datasets/) and 
extracted using the command `tar -xzf $filename`.

Datasets used: 

11mm Arrays

- [Human Glioblastoma 11mm](https://www.10xgenomics.com/datasets/human-brain-cancer-11-mm-capture-area-ffpe-2-standard)
- [Human Lung Cancer 11mm](https://www.10xgenomics.com/datasets/human-lung-cancer-11-mm-capture-area-ffpe-2-standard)
- [Human Ovarian Carcinoma 11mm](https://www.10xgenomics.com/datasets/human-ovarian-cancer-11-mm-capture-area-ffpe-2-standard)


6.5mm Arrays

- [Human Skin Melanoma 6.5mm](https://www.10xgenomics.com/datasets/human-melanoma-if-stained-ffpe-2-standard)
- [Human Lung Squamous Cell Carcinoma 6.5mm](https://www.10xgenomics.com/datasets/human-lung-cancer-ffpe-2-standard)
- [Human Ovarian Cancer 6.5mm](https://www.10xgenomics.com/datasets/human-ovarian-cancer-1-standard)


Since no full resolution images were available, we loaded the `detected_tissue_image.jpg` files for manual alignment in the 10X Loupe Browser version 9, instead.

After aligning the fiducial markers and selecting all spots, the `manual_alignment.png` images and the `${Slide_ID}.json` file was downloaded.

To approach the question of deriving spatially resolved spot positions, the downloaded JSON files were validated internally, as well as they were compared to the tissue position files for each sample.

---

# Define Settings and load functions

```{r settings}
# define directories
base_dir <- file.path("D:", "Visium_spatial")
out_dir <- file.path(base_dir, "results")
```

---

# Internal Validation of the 10X Loupe Browser generated Manual Alignment JSON file

Loading the Manual Alignment JSON file using `jsonlite::fromJSON()` returns a
list with geometry-related slots:

- `transform` provides a 3x3-transformation matrix.
- `oligo` includes a `data.frame` object with geometry information for 
each spot on the capture area.
- `fiducials` includes a `data.frame` object with geometry information 
for each spot of the fiducial frame and fiducial symbols.

The geometry information in `oligo` and `fiducials` comprises 
information for eight columns:

- slide geometry (`x`, `y`)
- spot positions within a orange-crate packing grid (`row`, `col`)
- image positions (`imageX`, `imageY`)
- `dia` with spot diameters (not in micrometer)
- whether the spot was selected to be in the `tissue`

---

## Fiducial Symbol Positions

For aligning the Fiducial symbols with their visual positions in the manual-alignment resp. detected-tissue image,  while `array_col` needs to be on the horizontal axis and increases from left to right, whereas `array_row` is represented by the reversed vertical axis and increases from top to bottom. 


| Array Position     | Image Position     |
| :----------------- | :----------------- |
| $row_{min}$        | Upper Image        |
| $row_{max}$        | Lower Image        |
| $col_{min}$        | Left Image         |
| $col_{max}$        | Right Image        |

Visual inspection of the detected-tissue images showed the same symbol arrangement across all six selected datasets. Consistent with the Visium slide convention placing the HOURGLASS fiducial marker in the upper-left corner, the four named fiducial symbols occurred at the same relative frame positions:

| Fiducial Symbol    | Image Position     |
| :----------------- | :----------------- |
| HOURGLASS          | Upper left         |
| TRIANGLE           | Lower left         |
| HEXOPEN            | Lower right        |
| HEXFILLED          | Upper right        |


The fiducial slide-coordinate ranges were also consistent within each array format, depending on the capture area size,
supporting a stable correspondence between the JSON fiducial coordinate system and the physical fiducial-frame layout.

```{r fiducial-symbols, echo=FALSE}
symbol_df <- utils::read.csv(file.path(out_dir, "symbol_df.csv"), row.names = 1)
corner_positions <- c("Upper_left", "Lower_left", "Lower_right", "Upper_right")
knitr::kable(symbol_df[, !colnames(symbol_df) %in% corner_positions], digits = 9, caption = "<span style='font-size: 22px; font-weight: bold;'>Fiducial Symbol Positions</span>")
```


get_ranges <- function(symbol_df,col){
  range_col <- strsplit(symbol_df[,col], split = "-")
  names(range_col) <- rownames(symbol_df)
  lapply(range_col, function(x){
    diff(as.numeric(x))})}
    
ranges_x <- get_ranges(symbol_df, "range_x")
ranges_y <- get_ranges(symbol_df, "range_y")
ranges_imageX <- get_ranges(symbol_df, "range_imageX")
ranges_imageY <- get_ranges(symbol_df, "range_imageY")

--- 

## Matrix models

The inverse of the forward ordinary least squares (OLS) transformation $solve(M) = M^{-1}$ 
was compared with a second OLS transformation fitted independently in the reverse direction $M^{-1}_{lm}$.
Because the two regressions minimize residuals in different dependent variables,  $solve(M)$ and $M^{-1}_{lm}$ 
are expected to be close but need not be numerically identical.

$$
M_{var \rightarrow pred}=
\begin{bmatrix} 
    a & b & c \\ 
    d & e & f \\
    0 & 0 & 1
\end{bmatrix} \approx M_{lm} =  \begin{bmatrix} 
    \text{pred}_x[var_x] &  \text{pred}_x[var_y] &  \text{pred}_x[Intercept] \\ 
     \text{pred}_y[var_y] &  \text{pred}_y[var_y] &  \text{pred}_y[Intercept]  \\
    0 & 0 & 1
\end{bmatrix}\\

\ \\ \ \\

\widehat{M^{-1}} = solve(M)  | \ det(M) \neq 0  \approx (\widehat{M})^{-1}= M^{-1}_{lm}  \\
\ \\ \ \\
M^{-1}_{pred \rightarrow var} = 
\frac{1}{ae - db}
\begin{bmatrix} 
    e & -b &  bf - ec \\ 
    -d & a &  dc - af \\
    0 & 0 & ae -db
\end{bmatrix} | \ det(M) \neq 0 \approx
M^{-1}_{lm} =
\begin{bmatrix} 
    \text{var}_x[pred_x] & \text{var}_x[pred_y] & \text{var}_x[intercept] \\ 
   \text{var}_y[pred_x] & \text{var}_y[pred_y] & \text{var}_y[intercept] \\ 
    0 & 0 & 1
\end{bmatrix}
$$

---

# Image-Slide JSON Transformation Matrices

Validation of the 3*3-transformation matrix from  Manual Alignment JSON file ensuring correct 
alignment between Loupe slide-layout coordinates (`x`,`y`) and Loupe image coordinates (`imageX`, `imageY`).

## Encoded transformation Steps

All transformation matrices in the investigated Loupe manual-alignment JSON files `transform` slot 
have the form of a two-dimensional similarity transformation with uniform scaling and rotation 
along with a translation enabled by the 3rd dimension (c,d):

$$
A =
\begin{bmatrix} 
    a & b & c \\ 
    -b & a & d \\
    0 & 0 & 1
\end{bmatrix}
$$

Overview of the transformation steps and the corresponding matrix values:

| Matrix values        | Transformation     |
| :------------------- | :----------------- |
| (c,d)                | Translation        |
| (a,b)                | Rotation           |
| ($\sqrt{a^2 + b^2}$) | Uniform Scaling    |


If $det(A)= a^2+b^2 \neq 0$, an inverse matrix $solve(A)= A^{-1}$ exists, with:

$$
A^{-1} =
\frac{1}{a^2 + b^2}
\begin{bmatrix} 
    a & -b &  bd - ca \\ 
    b & a &  -bc -ad \\
    0 & 0 & a^2 + b^2
\end{bmatrix} | \ det(A) = a^2+b^2 \neq 0
$$

---

1) Slide $\rightarrow$ image transformation

Does the transformation matrix in the JSON file encode slide $\rightarrow$ image transformation?

### Linear Models

For ensuring proper interpretation, all 3x3-transformation matrices $A$ from the `transform` slot 
were compared to the Slide $\rightarrow$ Image transformation matrix $A_{lm}$ obtained by the linear 
model coefficients for `A_imageX <- stats::coef(stats::lm(imageX ~ x+y, data = oligo))` and 
`A_imageY <- stats::coef(stats::lm(imageY ~ x+y, data = oligo))` with:

$$
A_{lm} =
\begin{bmatrix} 
    A_{imageX}[x] & A_{imageX}[y] &  A_{imageX}[intercept] \\ 
    A_{imageY}[x] & A_{imageY}[y] & A_{imageY}[intercept] \\
    0 & 0 & 1
\end{bmatrix}
$$

All 3x3-transformation matrices $A$ from the `transform` slot were compared to the results 
of the linear model (`lm_identical`) and saved to `slide2image_transformation_matrix_A.csv`:

```{r A-slide2image, echo=FALSE}
df_A <- utils::read.csv(file.path(out_dir, "slide2image_transformation_matrix_A.csv"))
knitr::kable(df_A, digits = 9, caption = "<span style='font-size: 22px; font-weight: bold;'>Transformation matrices for slide -> image conversion</span>")
```


### Maximum Errors

The maximum error related to the slide to image coordinate transformation
$\text{slide coordinates} \overset{A}{\rightarrow} \text{image coordinates}$ 
across all samples is very small:

```{r error-slide2image, echo=FALSE}
df_A <- utils::read.csv(file.path(out_dir, "slide2image_transformation_inaccuracy.csv"))
knitr::kable(df_A, digits = 9, caption = "<span style='font-size: 22px; font-weight: bold;'>Transformation inaccuracy for slide -> image conversion</span>")
```

This indicates a reliable transformation using the sample-specific 
transformation matrix $A$ for conversion of slide to image coordinates.

---

2) Image $\rightarrow$ slide transformation

Can the transformation be inverted to encode image $\rightarrow$ slide transformation?

### Linear Models

The Image $\rightarrow$ Slide transformation matrix $A^{-1}$ was primarily built using `solve(A)`.
Additionally, an inverse transformation matrix $A^{-1}_{lm}$ was built using the linear model coefficients
`A_x <- stats::coef(stats::lm(x ~ imageX + imageY, data = oligo))` and
`A_y <- stats::coef(stats::lm(y ~ imageX + imageY, data = oligo))` with:

$$
A^{-1}_{lm} =
\begin{bmatrix} 
    A_x[imageX] & A_x[imageY] & A_x[intercept] \\ 
    A_y[imageX] & A_y[imageY] & A_y[intercept] \\
    0 & 0 & 1
\end{bmatrix}
$$

All inverted 3x3-transformation matrices $A^{-1}$, calculated using `solve(A)` 
were compared with the results of the linear model (`lm_identical`) and saved 
as `image2slide_transformation_matrix_A_inv.csv`:

```{r A-inv-image2slide, echo=FALSE}
df_A_inv<- utils::read.csv(file.path(out_dir, "image2slide_transformation_matrix_A_inv.csv"))
knitr::kable(df_A_inv, digits = 9, caption = "<span style='font-size: 22px; font-weight: bold;'>Transformation matrices for image -> slide conversion</span>")
```


### Maximum Errors

Again, the maximum error related to the image to slide coordinate transformation
$\text{image coordinates} \overset{A^{-1}}{\rightarrow} \text{slide coordinates}$ 
across all samples is very small:

```{r error-image2slide, echo=FALSE}
df_A_inv <- utils::read.csv(file.path(out_dir, "image2slide_transformation_inaccuracy.csv"))
knitr::kable(df_A_inv, digits = 9, caption = "<span style='font-size: 22px; font-weight: bold;'>Transformation inaccuracy for image -> slide conversion</span>")
```

Therefore, the sample-specific inverse transformation matrix $A^{-1}$, can
be used for converting image to slide coordinates with very high precision.

---

### Summary

Altogether, the transformation matrix $A$ and its inverse $solve(A)=A^{-1}$ reproduce 
the JSON coordinates image and slide coordinates with very high numeric accuracy.
Both transformation matrices can be constructed from the linear model coefficients 
(`stats::coeff(stats::lm(model))`), independently.


$$
\text{slide coordinates} \overset{A}{\rightarrow} \text{image coordinates}\\
\text{image coordinates} \overset{A^{-1}}{\rightarrow} \text{slide coordinates}
$$


---

## Diemeter information

Can the diameter information be used for inferring spatial information?

For converting image $\rightarrow$ slide coordinates, the scaling factor is determined 
by $sf_{A^{-1}} = det(A)= \sqrt{a^2 + b^2}$

```{r slide-coordinate-df, echo=FALSE}
diameter_df <- utils::read.csv(file.path(out_dir,  "diameter_df.csv"), row.names = 1)
knitr::kable(diameter_df, digits = 9, caption = "<span style='font-size: 22px; font-weight: bold;'>Spot Diameter</span>")
```



---

## Slide Coordinates

Do the slide coordinate distances dx and dy encode spatial distances in micrometer, and do
the observed distances match any of the cases illustrated in [Reconstruction_models](https://Zockimonster.github.io/visiumSpatialSolutions/)? 

```{r slide-coordinate-df, echo=FALSE}
slide_df <- utils::read.csv(file.path(out_dir, "slide_coordinates.csv"), row.names = 1)
knitr::kable(slide_df, digits = 9, caption = "<span style='font-size: 22px; font-weight: bold;'>Slide Coordinates</span>")
```

For the observed slide coordinates (`x`,`y`) all fitted column increments (`dx`) equal 50 
and all fitted row increments (`dy`) 87 for the 6.5mm array and are close to 86.5 for the 11mm array.

Therefore, the column increments (`dx`) are consistent with the manufacturer's reported 100µm nearest neighbor distances $2dx = 100$,
and all observed ratios for the fitted slide coordinate increments are consistent with the expected ratio of column 
and row distances within a hexagonal grid as defined by case 3 in [Reconstruction_Models](https://Zockimonster.github.io/visiumSpatialSolutions/), with $(d_{diag} = 2d_{col} \wedge d_{diag} =\sqrt{dx^2 + dy^2}) \implies \frac{d_{col}}{d_{row}} = \frac{1}{\sqrt{3}} \approx 0.5773503$.

---

# Tissue position relationship

The number of spots in the JSON and tissue position file as well as 
the unique array coordinates match perfectly between both files:

```{r n-spots, echo=FALSE}
spots_df <- utils::read.csv(file.path(out_dir, "n_spots_combined.csv"))
knitr::kable(spots_df, digits = 9, caption = "<span style='font-size: 22px; font-weight: bold;'>Number of spots</span>")
```


---

## Corresponding Tissue Positions

To ensure the identical meaning of array coordinates across both files, we applied following check:

Manual Selection of identical spots based on `detected_tissue_image`
      $\downarrow$  
Combination of returned JSON file and tissue positions   
      $\downarrow$      
Do selected tissue spots match between JSON and tissue positions file?


Revealing perfect alignment of array coordinates for `spatial_Human_Ovarian_Cancer`:

```{r tp-small-array, echo=FALSE}
utils::read.csv(file.path(out_dir, "tissue_position_check_spatial_Human_Ovarian_Cancer.csv"))
```


as well as for `spatial_Human_Ovarian_Carcinoma`:
        
```{r tp-large-array, echo=FALSE}
utils::read.csv(file.path(out_dir, paste0("tissue_position_check_spatial_Human_Ovarian_Carcinoma.csv")))
```


--- 

## Relation between JSON and tissue position image coordinates


1) JSON $\rightarrow$ TP image coordinate Transformation

### Linear Models

The JSON $\rightarrow$ TP image transformation matrix $B$ was built using the linear model coefficients
`B_pxl_row <- stats::coef(stats::lm(pxl_row_in_fullres ~ imageX + imageY, data = tp_json))` and 
`B_pxl_col <- stats::coef(stats::lm(pxl_col_in_fullres ~ imageX + imageY, data = tp_json))` with:

$$
B_{\ \text{pixel}_{JSON} \rightarrow \ \text{pixel}_{TP}} =
\begin{bmatrix} 
    B_{pxl\_row}[imageX] & B_{pxl\_row}[imageY] &  B_{pxl\_row}[intercept] \\ 
     B_{pxl\_col}[imageX] & B_{pxl\_col}[imageY] & B_{pxl\_col}[intercept] \\
    0 & 0 & 1
\end{bmatrix} =
\begin{bmatrix} 
    a_{11} & a_{12} & b_1 \\ 
    a_{21} & a_{22} & b_2 \\
    0 & 0 & 1
\end{bmatrix}

\ \\ \ \\
\text{pixel}_{JSON} \overset{B}{\rightarrow} \text{pixel}_{TP}
$$

All 3x3 transformation matrices $B$ for JSON $\rightarrow$ TP image coordinate 
conversion derived by the linear models were saved to `JSON2TP_transformation_matrix_B.csv`:

```{r B-JSON2TP, echo=FALSE}
df_B <- utils::read.csv(file.path(out_dir, "JSON2TP_transformation_matrix_B.csv"))
knitr::kable(df_B, digits = 9, caption = "<span style='font-size: 22px; font-weight: bold;'>Transformation matrices for JSON -> TP image conversion</span>")
```

 df_B
                            Sample_Name Array_Size  row          col1          col2         col3
1                  spatial_Glioblastoma       11mm row1   0.075995047 -16.758784115 49017.546288
2                  spatial_Glioblastoma       11mm row2  16.758785381   0.075993333  9696.957221
3                  spatial_Glioblastoma       11mm row3   0.000000000   0.000000000     1.000000
4          spatial_Human_Ovarian_Cancer      6.5mm row1   0.002485077   6.889392405    -4.624027
5          spatial_Human_Ovarian_Cancer      6.5mm row2   6.889395144  -0.002482931    -5.069893
6          spatial_Human_Ovarian_Cancer      6.5mm row3   0.000000000   0.000000000     1.000000
7       spatial_Human_Ovarian_Carcinoma       11mm row1  -0.201207646  16.938440506 -5449.799843
8       spatial_Human_Ovarian_Carcinoma       11mm row2 -16.938440791  -0.201207658 38458.905706
9       spatial_Human_Ovarian_Carcinoma       11mm row3   0.000000000   0.000000000     1.000000
10                  spatial_Lung_Cancer       11mm row1   0.251682471 -16.910645246 38951.964921
11                  spatial_Lung_Cancer       11mm row2  16.910644475   0.251682260 -4664.084691
12                  spatial_Lung_Cancer       11mm row3   0.000000000   0.000000000     1.000000
13 spatial_Lung_Squamous_Cell_Carcinoma      6.5mm row1   0.127001375 -16.952528955 39713.897893
14 spatial_Lung_Squamous_Cell_Carcinoma      6.5mm row2  16.952525994   0.127001875  -186.319012
15 spatial_Lung_Squamous_Cell_Carcinoma      6.5mm row3   0.000000000   0.000000000     1.000000
16                spatial_Skin_Melanoma      6.5mm row1  -0.026575963  14.117722945 -8989.300573
17                spatial_Skin_Melanoma      6.5mm row2 -14.117712630  -0.026565339 24172.213430
18                spatial_Skin_Melanoma      6.5mm row3   0.000000000   0.000000000     1.000000


### Maximum Errors

The maximum error related to the JSON to TP image coordinate transformation
$\text{pixel}_{JSON} \overset{B}{\rightarrow} \text{pixel}_{TP}$ 
across all samples is relatively small:

```{r error-JSON2TP, echo=FALSE}
df_B <- utils::read.csv(file.path(out_dir, "JSON2TP_transformation_inaccuracy.csv"))
knitr::kable(df_B, digits = 9, caption = "<span style='font-size: 22px; font-weight: bold;'>Transformation inaccuracy for JSON -> TP image conversion</span>")
```

df_B
                           Sample_Name Array_Size max_abs_pxl_row max_abs_pxl_col
1                 spatial_Glioblastoma       11mm       0.5025436       0.5015630
2         spatial_Human_Ovarian_Cancer      6.5mm       0.5003504       0.5019114
3      spatial_Human_Ovarian_Carcinoma       11mm       0.5004045       0.5028164
4                  spatial_Lung_Cancer       11mm       0.5021980       0.5004690
5 spatial_Lung_Squamous_Cell_Carcinoma      6.5mm       0.5038978       0.5012755
6                spatial_Skin_Melanoma      6.5mm       0.4863538       0.5020253


This indicates a sufficiently reliable transformation using the sample-specific 
transformation matrix $B$ for conversion of JSON to TP image coordinates.

---

2) TP $\rightarrow$ JSON image coordinate Transformation

If $det(B)= a_{11}a_{22} - a_{12}a_{21}  \neq 0$, an inverse matrix $solve(B)=B^{-1}$ exists, with:

$$
B^{-1}_{\ \text{pixel}_{TP} \rightarrow \ \text{pixel}_{JSON}} =
\frac{1}{a_{11}a_{22} - a_{12}a_{21}}
\begin{bmatrix} 
    a_{22} & -a_{12} &  a_{12}b_2 - a_{22}b_1 \\ 
    -a_{21} & a_{11} &  a_{21}b_1 -a_{11}b_2 \\
    0 & 0 & a_{11}a_{22} - a_{12}a_{21}
\end{bmatrix} | \ det(B) = a_{11}a_{22} - a_{12}a_{21} \neq 0

\ \\ \ \\ \ \\
\text{pixel}_{TP} \overset{B^{-1}}{\rightarrow} \text{pixel}_{JSON} \\
$$

### Linear Models

The TP $\rightarrow$ JSON image transformation matrix $B^{-1}$ was primarily built using `solve(B)`.
Additionally, an inverse transformation matrix $B^{-1}_{lm}$ was built using the linear model coefficients
`B_imageX <- stats::coef(stats::lm(imageX ~ pxl_row_in_fullres + pxl_col_in_fullres, data = tp_json))` and 
`B_imageY <- stats::coef(stats::lm(imageY ~ pxl_row_in_fullres + pxl_col_in_fullres, data = tp_json))` with:

$$
B^{-1}_{lm} =
\begin{bmatrix} 
    B_{imageX}[pxl\_row\_in\_fullres] & B_{imageX}[pxl\_col\_in\_fullres] &  B_{imageX}[intercept] \\ 
    B_{imageY}[pxl\_row\_in\_fullres] & B_{imageY}[pxl\_col\_in\_fullres] & B_{imageY}[intercept] \\
    0 & 0 & 1
\end{bmatrix}
$$


All inverted 3x3-transformation matrices $B^{-1}$, calculated using `solve(B)` 
were compared with the results of the linear model (`lm_identical`) and saved 
as `TP2JSON_transformation_matrix_B_inv.csv`:

```{r B-inv-TP2JSON, echo=FALSE}
df_B_inv<- utils::read.csv(file.path(out_dir, "TP2JSON_transformation_matrix_B_inv.csv"))
knitr::kable(df_B_inv, digits = 9, caption = "<span style='font-size: 22px; font-weight: bold;'>Transformation matrices for TP -> JSON image conversion</span>")
```

df_B_inv
                            Sample_Name Array_Size                         lm_identical  row          col1          col2         col3
1                  spatial_Glioblastoma       11mm                                 TRUE row1  2.705712e-04  5.966897e-02 -591.8701702
2                  spatial_Glioblastoma       11mm                                 TRUE row2 -5.966897e-02  2.705773e-04 2922.2028728
3                  spatial_Glioblastoma       11mm                                 TRUE row3  0.000000e+00  0.000000e+00    1.0000000
4          spatial_Human_Ovarian_Cancer      6.5mm Mean relative difference: 1.3212e-05 row1  5.231216e-05  1.451506e-01    0.7361399
5          spatial_Human_Ovarian_Cancer      6.5mm Mean relative difference: 1.3212e-05 row2  1.451507e-01 -5.235735e-05    0.6709151
6          spatial_Human_Ovarian_Cancer      6.5mm Mean relative difference: 1.3212e-05 row3  0.000000e+00  0.000000e+00    1.0000000
7       spatial_Human_Ovarian_Carcinoma       11mm                                 TRUE row1 -7.011911e-04 -5.902898e-02 2266.3686981
8       spatial_Human_Ovarian_Carcinoma       11mm                                 TRUE row2  5.902898e-02 -7.011910e-04  348.6631814
9       spatial_Human_Ovarian_Carcinoma       11mm                                 TRUE row3  0.000000e+00  0.000000e+00    1.0000000
10                  spatial_Lung_Cancer       11mm                                 TRUE row1  8.799056e-04  5.912126e-02  241.4724938
11                  spatial_Lung_Cancer       11mm                                 TRUE row2 -5.912125e-02  8.799063e-04 2306.9929472
12                  spatial_Lung_Cancer       11mm                                 TRUE row3  0.000000e+00  0.000000e+00    1.0000000
13 spatial_Lung_Squamous_Cell_Carcinoma      6.5mm                                 TRUE row1  4.418927e-04  5.898495e-02   -6.5592649
14 spatial_Lung_Squamous_Cell_Carcinoma      6.5mm                                 TRUE row2 -5.898494e-02  4.418910e-04 2342.6041602
15 spatial_Lung_Squamous_Cell_Carcinoma      6.5mm                                 TRUE row3  0.000000e+00  0.000000e+00    1.0000000
16                spatial_Skin_Melanoma      6.5mm                                 TRUE row1 -1.332861e-04 -7.083275e-02 1710.9862851
17                spatial_Skin_Melanoma      6.5mm                                 TRUE row2  7.083270e-02 -1.333394e-04  639.9595541
18                spatial_Skin_Melanoma      6.5mm                                 TRUE row3  0.000000e+00  0.000000e+00    1.0000000


### Maximum Errors

The maximum error related to the TP to JSON image coordinate transformation
$\text{pixel}_{TP} \overset{B^{-1}}{\rightarrow} \text{pixel}_{JSON}$ 
across all samples is very small:

```{r error-TP2JSON, echo=FALSE}
df_B_inv <- utils::read.csv(file.path(out_dir, "TP2JSON_transformation_inaccuracy.csv"))
knitr::kable(df_B_inv, digits = 9, caption = "<span style='font-size: 22px; font-weight: bold;'>Transformation inaccuracy for TP -> JSON image conversion</span>")
```

df_B_inv
                           Sample_Name Array_Size max_abs_imageX max_abs_imageY
1                 spatial_Glioblastoma       11mm     0.03000334     0.03008183
2         spatial_Human_Ovarian_Cancer      6.5mm     0.07283828     0.07263399
3      spatial_Human_Ovarian_Carcinoma       11mm     0.02982102     0.02984054
4                  spatial_Lung_Cancer       11mm     0.02992354     0.02993835
5 spatial_Lung_Squamous_Cell_Carcinoma      6.5mm     0.02957716     0.02975440
6                spatial_Skin_Melanoma      6.5mm     0.03559712     0.03447388

Therefore, the sample-specific inverse transformation matrix $B^{-1}$, can
be used for converting TP to JSON image coordinates with very high precision.

---

### Summary

Altogether, the transformation matrix $B$ and its inverse $solve(B)=B^{-1}$ reproduce 
the JSON and TP image coordinates with high numeric accuracy,
The independently fitted reverse linear transformation (`stats::coeff(stats::lm(model))`)
produced nearly identical results across all investigated datasets.

All JSON and tissue positions can be converted using affine transformation, though
the exact conversion appears to be sample-specific, including all kinds of affine 
transformation procedures.


## Affine Transformation procedures

| Transformation                  | Linear model values               |
| :------------------------------ | :---------------------------------|
| Translation                     | Intercept differs                 |
| Reflection                      | Negative Slope                    |
| Scaling / Cropping / Resizing   | Absolute Slope != 1               |
| Cross-axis contribution         | Axis-specific transformation      |


---

# Reproducibility

## Script availability

The calculations, tables and figures shown on this page are generated from:

- [`code_spatial_alignment.R`](https://Zockimonster.github.io/visiumSpatialSolutions/demo_spatial/code_spatial_alignment.R)
- [`functions_spatial_alignment.R`](https://Zockimonster.github.io/visiumSpatialSolutions/demo_spatial/functions_spatial_alignment.R)


Specific results comprising all related models and measurements are also available in:

- [results](https://Zockimonster.github.io/visiumSpatialSolutions/demo_spatial/results/)

## Session Info

```{r}
sessionInfo()
```



